---
title: "CRB Variance"
author: "Brian"
date: "6/14/2022"
output:
  html_document:
    toc: true
    toc_float: true
editor_options: 
  chunk_output_type: console
---

# Disclaimer 

*This work is very preliminary as I get back into the coding swing of things. Data wrangling and figure generation will be done via R, but the rest of the project will be done using good ol' microsoft products. This is just an entry point into data crunching and should by no means be considered a final product.*

# Steamboat Tower SNOTEL

```{r libraries, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(snotelr)
library(riem)
library(dplyr)
library(ggplot2)
library(tidyverse)
library(dataRetrieval)
library(lubridate)
library(sf)
library(ggthemes)
library(xts)
library(dygraphs)
library(scales)
library(openair)
library(plotly)
library(SciViews)
knitr::opts_chunk$set(message = F, 
                      warning = F,
                      cache = T)
knitr::opts_chunk$set(echo = TRUE)
library(trend)
library(nhdplusTools)
library(lfstat)
library(ggpubr)
library(kableExtra)

#Stats
library(forcats)
library(stringr)
library(trend)

# COOP/ Did not work
#library(rnoaa)
```

## Raw data download

### Steamboat Tower (825) SNOTEL data

```{r read in and download, echo=TRUE}
# The fucntion snotel_download evidently downloads everything in metric units. 

snotel_825_tower <- snotel_download(site_id = 825, path = tempdir('../data'), internal = TRUE)

write.csv(snotel_825_tower,"C:/Users/13074/Documents/ESS580/thesis_project/thesis_research/data_raw/snotel_825.csv", row.names = FALSE) #write in the raw data

```

### Data Cleaning

```{r clean}
head(snotel_825_tower) # check the date, usually a character.  

snotel_825_tower$Date <- as.Date(snotel_825_tower$date) #change date from character to date format, capitalize to work with Water year functon from NWIS.

snotel_825_clean <- snotel_825_tower %>% # filter for the timeframe
  filter(Date >= "1978-10-01" & Date <= "2021-09-30") %>%
  filter(temperature_mean >= -30 & temperature_mean <= 20) %>% # removing outliers   
  addWaterYear() %>% 
  mutate(daymonth = format(as.Date(Date), "%d-%m")) %>% 
  na.omit()
```

```{r 825, eval=FALSE, include=FALSE}
snotel_825$Date <- as.Date(snotel_825$date) #change date from character to date format, capitalize to work with Water year functon from NWIS.

snotel_825_clean <- snotel_825 %>% # filter for the timeframe
  filter(Date >= "1978-10-01" & Date <= "2021-09-31") %>%
  filter(temperature_mean >= -30 & temperature_mean <= 20) %>% # removing outliers   
  addWaterYear() %>% 
  mutate(daymonth = format(as.Date(Date), "%d-%m")) %>%
  mutate(DOY = yday((Date))) %>%
  na.omit()
```

##### adding water day using difftime (SUPER COOL. example from [this](https://stackoverflow.com/questions/48123049/create-day-index-based-on-water-year))

```{r clean water day & save}

#adding water day using difftime (SUPER COOL. example from [this](https://stackoverflow.com/questions/48123049/create-day-index-based-on-water-year))

snotel_825_clean <- snotel_825_clean %>% 
  group_by(waterYear)%>% 
  mutate(waterDay = (as.integer(difftime(Date, ymd(paste0(waterYear - 1 ,'-09-30')), units = "days"))))


write.csv(snotel_825_clean,"C:/Users/13074/Documents/ESS580/thesis_project/thesis_research/data_clean/snotel_825_clean.csv", row.names = FALSE)

```

### Figure check

```{r 825 simple plot}
ggplot(snotel_825_clean, aes(x = Date, y = temperature_mean, filter(waterYear = 1990))) + #this filter didn't work.....
  geom_point() + #lwd = 2) +
  theme_few() +
  geom_smooth(method = "lm", se=FALSE) +
  ylab('Daily temperature (°C)') + 
  xlab('Date')

#Check for outliers....

#dygraph

temp_xts <- xts(snotel_825_clean$temperature_mean, order.by = snotel_825_clean$Date)

dygraph(temp_xts) %>%
  dyAxis("y", label = "Daily temperature (°C)") 


```

## Detrending Data 

```{r detrending data}
#SF figured out the yearly average by water year

#average water year temperature

yearly_wy_aver <- snotel_825_clean %>% 
  group_by(waterYear) %>% 
  mutate(aver_ann_temp = mean(temperature_mean))

```


```{r detrending data2}
#Average temperature by day for all water years:

daily_wy_aver <- yearly_wy_aver %>% 
  group_by(daymonth) %>% 
  mutate(aver_day_temp = mean(aver_ann_temp))

#average mean temperature by day for the period of record:

daily_wy_aver <- daily_wy_aver %>% 
  group_by(daymonth) %>% 
  mutate(all_ave_temp = mean(daily_wy_aver$aver_day_temp))

#str(daily_wy_aver)

```

```{r detrend figures, eval=FALSE, include=FALSE}
#making day-month numeric

head(daily_wy_aver)

# try to show all years as means. Uses Jan 1 as day 1. 
daily_wy_aver2 <- daily_wy_aver %>% 
  #filter(waterYear == "1987" | waterYear == "2021") %>%
  select(DOY, daymonth, aver_day_temp) %>%
  distinct(daymonth, .keep_all = TRUE) %>% # had to use daymonth to keep it distinct
  mutate(date = as.Date(DOY)) #This worked, but gave a year of 1970?

daily_wy_aver2$aver_day_temp <- signif(daily_wy_aver2$aver_day_temp,2) #reduce the sig figs

daily_wy_aver2$DOYwy <- 1:nrow(daily_wy_aver2) #adding a number as a wateryear day of year.  

### ### ###

#NONE OF THIS WAS RIGHT- I WANTED SIMPLY THE AVERAGE TEMPERATURE BY waterDay ACROSS ALL YEARS! 
# try to show all years as means. 
daily_wy_aver2 <- daily_wy_aver %>% 
  #filter(waterYear == "1987" | waterYear == "2021") %>%
  select(waterDay, daymonth, aver_day_temp) %>%
  distinct(daymonth, .keep_all = TRUE) #%>% # had to use daymonth to keep it distinct
  #mutate(date = as.Date(DOY)) #This worked, but gave a year of 1970?
  #NO NEED TO DO THIS, HAVE waterDay NOW.

daily_wy_aver2$aver_day_temp <- signif(daily_wy_aver2$aver_day_temp,2) #reduce the sig figs



```

```{r Figure all year average temp}

# try to show all years as means. 
daily_wy_aver2 <- daily_wy_aver %>% 
  #filter(waterYear == "1987" | waterYear == "2021") %>%
  group_by(waterDay) %>%
  mutate(date_temp = mean(temperature_mean))
  

daily_wy_aver2$date_temp <- signif(daily_wy_aver2$date_temp,3) #reduce the sig figs



ggplot(daily_wy_aver2, aes(x = waterDay, y = date_temp))+#, color = waterYear)) +
  geom_line(size= 0.7) +
  #geom_line(aes) +
  theme_few() +
  #geom_smooth(method = "lm", se=FALSE) +
  ylab('Average Daily temperature (°C)') + 
  xlab('Day of water year')

temp_xts_2 <- xts(daily_wy_aver2$date_temp, order.by = as.Date(daily_wy_aver2$waterDay))

dygraph(temp_xts_2) %>%
  dyAxis("y", label = "(°C)") 

```
**Day of year average temperature for the 1987-2021 period of record for Tower SNOTEL site.**


# The temperature plateaus in summer
From 266 to 340, there are strange temperature plateaus in the figures. 

```{r}

# try to show all years as means. 
plateau_daily_wy_aver <- daily_wy_aver %>% 
  #filter(waterYear == "1987" | waterYear == "2021") %>%
  filter(waterDay == 266) %>% 
  group_by(waterDay) %>% 
  mutate(date_temp = mean(temperature_mean)) %>% 
  as.tibble()

head(plateau_daily_wy_aver$date_temp, 1)  

```

Expanding significant figures fixed it.

## All years vs 1987 & 2021

```{r all year average join to 1987 & 2021 WY}



# 87 and 21 only. Not useful
daily_wy_aver3 <- daily_wy_aver2 %>% 
  filter(waterYear == "1987" | waterYear == "2021") %>%
  select(waterDay, waterYear, temperature_mean, daymonth, date_temp) %>%
  group_by(daymonth)

# ggplot(daily_wy_aver3, aes(x = waterDay, y = date_temp))+#, color = waterYear)) +
#   geom_line(size= 0.7) +
#   geom_line(aes(y = temperature_mean, color = waterYear)) +
#   theme_few() +
#   #geom_smooth(method = "lm", se=FALSE) +
#   #scale_colour_identity() +
#   scale_color_manual(name = "Water Year", values = c("blue", "red")) 
#   ylab('Daily temperature (°C)') + 
#   xlab('Day of water year')

ggplot(daily_wy_aver3) + #, color = waterYear)) +
  geom_line(aes(x = waterDay, y = date_temp), group = 1, size= 0.7) +
  geom_line(aes(x = waterDay, y = temperature_mean, group = 1, colour = waterYear)) +
  theme_few() +
  #geom_smooth(method = "lm", se=FALSE) +
  #scale_colour_identity() +
  
  #scale_color_manual(name = "Water Year") 
  ylab('Daily temperature (°C)') + 
  xlab('Day of water year')
  
# ??geom_line  
#   
# scale_color_manual(name = "Colors", 
#                      values = c("a" = "blue", "b" = "red")  
#   
# color_group <- c("blue","black")
#   
# ?scale_discrete_identity()

```

**Day of year average temperature and daily 1987 & 2021 temperatures for the 1987-2021 period of record for Tower SNOTEL site.** *legend is bad.*



### Calculating variance

```{r Calculating residuals}
#Calculating residuals, WY 1987 (first full year on record for 825)

residuals_1987 <- daily_wy_aver %>% 
  group_by(waterYear) %>% 
  filter(waterYear == 1987) %>% 
  mutate(residual = (aver_ann_temp-all_ave_temp)+temperature_mean-aver_day_temp) %>% 
  mutate(deviation = abs(residual-lag(residual)))

residuals_2021 <- daily_wy_aver %>% 
  group_by(waterYear) %>% 
  filter(waterYear == 2021) %>% 
  mutate(residual = (aver_ann_temp-all_ave_temp)+temperature_mean-aver_day_temp) %>% 
  mutate(deviation = abs(residual-lag(residual)))


```

## Figures 

### 1987

```{r residuals 1987}

ggplot(residuals_1987, aes(residual)) +
  geom_histogram(alpha = 0.5, color = "blue", bins = 15, binwidth = 2) +
  theme_base() +
  xlab("Temperature (°C) Above or Below Mean") +
  ylab("Count")

```

**Distribution of daily temperature variance for 1987 compared to the 1987-2021 period of record.**


```{r deviations 1987}

ggplot(residuals_1987, aes(deviation)) +
geom_histogram(alpha = 0.5, color = "blue", bins = 15, binwidth = 1) +
  scale_x_continuous(breaks = seq(0, 15, by=5)) +
  theme_base() +
  xlab("Diurnal Temperature Fluctuation (°C)") +
  ylab("Count")

```

**Distribution of 1987 diurnal temperature fluctuation from the 1987-2021 period of record.**

```{r percent of year 1987}

sum(residuals_1987$deviation, na.rm = TRUE)

865.0682

residuals_1987 <- residuals_1987 %>% 
  mutate(prcnt = deviation/865.0682*100)

sum(residuals_1987$prcnt, na.rm = TRUE)

ggplot(residuals_1987, aes(prcnt)) +
geom_histogram(alpha = 0.5, color = "blue", bins = 15, binwidth = 0.1) +
  #scale_x_continuous(breaks = seq(0, 15, by=5)) +
  theme_base() +
  xlab("Diurnal Temperature Fluctuation (°C)") +
  ylab("Percent of Year")


ggplot(residuals_1987) + #, color = waterYear)) +
  geom_jitter(aes(x = deviation, y = prcnt), width = 1, size=1) +
  theme_few() +
  #geom_smooth(method = "lm", se=FALSE) +
  #scale_colour_identity() +
  #scale_color_manual(name = "Water Year") 
  ylab('Percent of Year') + 
  xlab('Deviation (°C)')

```

**Percent of year 1987 for the deviation. This... doesn't look right.**

### 2021

```{r residuals 2021}

ggplot(residuals_2021, aes(residual)) +
  geom_histogram(alpha = 0.5, color = "blue", binwidth = 2) +
  theme_base() +
  xlab("Temperature (°C) Above or Below Mean") +
  ylab("Count")

```

**Distribution of daily temperature variance for 2021 compared to the 1987-2021 period of record.**

```{r deviations 2021}

ggplot(residuals_2021, aes(deviation)) +
  geom_histogram(alpha = 0.5, color = "blue", bins = 15, binwidth = 1) +
  scale_x_continuous(breaks = seq(0, 15, by=5)) +
  theme_base() +
  xlab("Diurnal Temperature Fluctuation (°C)") +
  ylab("Count")
# 
# ggplot(residuals_2021, aes(deviation)) +
# geom_histogram(binwidth=1,
#                  center = 6,
#                  aes(col=I("blue"))) +
#   scale_x_continuous(breaks = seq(0, 15, by=5))
#   scale_x_continuous(breaks=seq(1,max(residuals_2021$deviation) + 1, by = 2))

```

**Distribution of 2021 diurnal temperature fluctuation from the 1987-2021 period of record.** *Scale not the same as 1987*

```{r percent of year 2021}

sum(residuals_2021$deviation, na.rm = TRUE)

# 832.3876

residuals_2021 <- residuals_2021 %>% 
  mutate(prcnt = deviation/832.3876*100)

sum(residuals_2021$prcnt, na.rm = TRUE)

ggplot(residuals_2021, aes(prcnt)) +
geom_histogram(alpha = 0.5, color = "blue", bins = 15, binwidth = 0.1) +
  #scale_x_continuous(breaks = seq(0, 15, by=5)) +
  theme_base() +
  xlab("Diurnal Temperature Fluctuation (°C)") +
  ylab("Percent of Year")

# This looks weird. Maybe a scatter plot is more appropriate.

ggplot(residuals_2021, aes(prcnt)) +
geom_histogram(alpha = 0.5, color = "blue", bins = 15, binwidth = 0.1) +
  #scale_x_continuous(breaks = seq(0, 15, by=5)) +
  theme_base() +
  xlab("Diurnal Temperature Fluctuation (°C)") +
  ylab("Percent of Year")

ggplot(residuals_2021) + #, color = waterYear)) +
  geom_jitter(aes(x = deviation, y = prcnt), width = 1, size=1) +
  theme_few() +
  #geom_smooth(method = "lm", se=FALSE) +
  #scale_colour_identity() +
  #scale_color_manual(name = "Water Year") 
  ylab('Percent of Year') + 
  xlab('Deviation (°C)')

```

**Percent of year 2021 for the deviaation. This... doesn't look right.**


# Mann-Kendall & Sen's Slope

### Mann Kendal for Tower site daily temperatures

```{r MK Tower}

# excluding dates where no temperature data were recorded. 

str(daily_wy_aver)

daily_wy_aver4 <- daily_wy_aver2 %>% 
  filter(Date > "1986-07-17")

tower_mk <- mk.test(daily_wy_aver4$temperature_mean)
print(tower_mk)

```

### Sen's Slope for Tower site daily temperatures

```{r Sens Tower}

tower_sens <- sens.slope(daily_wy_aver4$temperature_mean)

print(tower_sens)

```

# SNOTEL Site corrections

*"The Morrissey method also shows more variations in temperature trends between sites than the Oyler method, which dampens site-to-site variability in temperature trends.* 


# Next steps / Further questions

### SNOTEL

Need to investigate site specific issues related to disturbances (general SNOTEL issues). *Unsure of methods*

~~Having issues with day-of-year and water year interactions. This keeps me from showing mean trendlines over the whole period of record as R does not want to recognize the x axis as anything other than numerical.~~ *Resolved*  

Would using a pivot table be better when displaying temperature for 240 sites? (What data management is best for that much data?) *Doesn't look like it's needed currently*

How to include oscillation influence into the data? 

### Beyond SNOTEL

What other data / site types do I want to include? 

### Coding

I am sure Matt would tell me to automate this. Not quite sure how, but it seems likely I could find a list of SNOTEL sites and using spatial data, exclude sites not within the CRB boundary. From there, I could write a for loop & function to use the snotelr function to download data. *Jack Reuland helped with this idea below. Now what to do with the large dataframe....*

# All stations:

Per Jack's recommendations, BUT, while there are 240 stations, I could only find 166 while clipping the NRCS SNOTEL points and a CRB polygon from CAPGISadmin. The NRCS stated the sites given were active, so perhaps there are 74 inactive sites?

```{r read in SNOTEL station site_ids from ArcGIS, echo=TRUE}

TableOfSNOTEL_StudyArea <- read.csv("C:/Users/13074/Desktop/MS_Watershed_Science/Literature/SNOTEL_sites_1_2.csv", header = TRUE)

studyarea_site_id <- as.vector(TableOfSNOTEL_StudyArea$site_id)

SNOTEL_StudyArea <- snotel_download(site_id = c(studyarea_site_id), internal = TRUE)

write.csv(SNOTEL_StudyArea,"C:/Users/13074/Documents/ESS580/thesis_project/thesis_research/data_raw/SNOTEL_StudyArea.csv", row.names = FALSE) #write in the raw data

```

```{r summary StudyArea}

head(SNOTEL_StudyArea)

```

# Coop Data

NOAA/NWS Cooperative Observer Network

Two stations near steamboat, according to this [site](https://wrcc.dri.edu/inventory/coop05.html). 
057936 02 SBTC2 UNITED STATES CO ROUTT +7 STEAMBOAT SPRINGS 40 30 17 -106 51 58  6636
057942 02 SSPC2 UNITED STATES CO ROUTT +7 STEAMBOAT SPRINGS 1 W 40 29 00 -106 51 00 6700

Downloading from the [Iowa Environmental Mesonet](https://mesonet.agron.iastate.edu/request/coop/fe.phtml?network=COCLIMATE). NOAA package not working, maybe due to user error. 

```{r Read in Steamboat COOP, echo=TRUE}
# coops_search(station_name = 057936,product = "air_temperature") # Handy R package did not work.

coop_steamboat <- read.csv("C:/Users/13074/Documents/ESS580/thesis_project/thesis_research/data_raw/steamboat_nwscoop_1900_2022.csv", header = TRUE)
```


```{r xts in Steamboat COOP}

coop_steamboat$Date <- mdy(coop_steamboat$day) #change date from character to date format, capitalize to work with Water year functon from NWIS.

coop_sb_cleanish <- coop_steamboat %>% 
  addWaterYear() %>% 
  mutate(daymonth = format(as.Date(Date), "%d-%m")) %>% 
  na.omit()

#str(coop_steamboat)

sb_coop_temp_xts <- xts(coop_sb_cleanish$avg_T_c, order.by = coop_sb_cleanish$Date)

dygraph(sb_coop_temp_xts) %>%
  dyAxis("y", label = "Daily temperature (°C)") 


```





# Notes from SF, 22/6:

Some initial thoughts: 1) I assume that frequency is the raw count, i.e., number of days? At some point, a % of the year may be better; 2) COOP data, such as at Steamboat, should have a much longer period of record if you want to examine change over time - setting up the analysis is a good start; 3) there is an inconsistency with the SNOTEL data (see https://doi.org/10.1029/2019WR025921); 4) we (https://doi.org/10.1029/2002WR001512, etc.) used 240 stations (metadata at https://doi.org/10.1029/2009WR007835), and I know that are a few are inactive, but likely not 74 - some are outside of the basin (https://agupubs.onlinelibrary.wiley.com/cms/asset/2af0fc20-a948-455d-bf0e-8f6e7dcc7ad9/wrcr12217-fig-0001.png) but still representative (a few can be removed, in hindsight, like the ones in the Laramie Range).

# Notes, Monday 27/6

Figured out the water day issue and applied it to the Day of water year figures (this was an issue that I spent waaaaaay too much time on). Went the wrong way with pivot tables, day of year, and using average annual temp for the mean aily figure.  Going to use Excel to check if correct. Goal for Tuesday: Determine daily percentages.

# Notes, Thursday 30/6

Need to discuss methods of identifying temperature shifts due to sensor changes.

Not sure how to show percent of days for data.

# *Corrected Variance* 19/7

The above chunks of code may be incorrect; instead of subtracting the mean temperature for the period of record from the mean annual temperature for each water year, I should switch that and subtract the mean annual temperature for each water year from mean temperature for the period of record to find the residuals:

(all_ave_temp - aver_ann_temp) + temperature_mean - aver_day_temp
The average of the residuals should be zero.

```{r corrected variance and derivation}

residuals_1987_2 <- daily_wy_aver %>% 
  group_by(waterYear) %>% 
  filter(waterYear == 1987) %>% 
  mutate(residual = (all_ave_temp-aver_ann_temp)+temperature_mean-aver_day_temp) %>% 
  mutate(deviation = abs(residual-lag(residual)))

mean(residuals_1987_2$residual)

residuals_2021_2 <- daily_wy_aver %>% 
  group_by(waterYear) %>% 
  filter(waterYear == 2021) %>% 
  mutate(residual = (all_ave_temp-aver_ann_temp)+temperature_mean-aver_day_temp) %>% 
  mutate(deviation = abs(residual-lag(residual)))

mean(residuals_2021_2$residual)

```

Based on the calculated means, this looks correct.

### 1987

```{r residuals 2 1987}

ggplot(residuals_1987_2, aes(residual)) +
  geom_histogram(alpha = 0.5, color = "blue", bins = 15, binwidth = 2) +
  theme_base() +
  xlab("Temperature (°C) Above or Below Mean") +
  ylab("Count")

```

**Distribution of daily temperature variance for 1987 compared to the 1987-2021 period of record.**


```{r deviations 2 1987}

ggplot(residuals_1987_2, aes(deviation)) +
geom_histogram(alpha = 0.5, color = "blue", bins = 15, binwidth = 1) +
  scale_x_continuous(breaks = seq(0, 15, by=5)) +
  theme_base() +
  xlab("Diurnal Temperature Fluctuation (°C)") +
  ylab("Count")

```

**Distribution of 1987 diurnal temperature fluctuation from the 1987-2021 period of record.**

### 2021

```{r residuals 2 2021}

ggplot(residuals_2021_2, aes(residual)) +
  geom_histogram(alpha = 0.5, color = "blue", binwidth = 2) +
  theme_base() +
  xlab("Temperature (°C) Above or Below Mean") +
  ylab("Count")

```

**Distribution of daily temperature variance for 2021 compared to the 1987-2021 period of record.**

```{r deviations 2 2021}

ggplot(residuals_2021_2, aes(deviation)) +
  geom_histogram(alpha = 0.5, color = "blue", bins = 15, binwidth = 1) +
  scale_x_continuous(breaks = seq(0, 15, by=5)) +
  theme_base() +
  xlab("Diurnal Temperature Fluctuation (°C)") +
  ylab("Count")
# 
# ggplot(residuals_2021, aes(deviation)) +
# geom_histogram(binwidth=1,
#                  center = 6,
#                  aes(col=I("blue"))) +
#   scale_x_continuous(breaks = seq(0, 15, by=5))
#   scale_x_continuous(breaks=seq(1,max(residuals_2021$deviation) + 1, by = 2))

```

**Distribution of 2021 diurnal temperature fluctuation from the 1987-2021 period of record.** *Scale not the same as 1987*

# Standard Deviation 

To figure out the standard deviation for each year, I want the "residual" for each daily value. 

The standard deviation will be the daily residual minus the mean of the residuals by water year, summed and squared, then divided by the number of observations minus one. The square root of the resulting value of which is thus the standard deviation for the water year. 


```{r sd}
standard_dev <- daily_wy_aver %>% 
  group_by(waterYear) %>% 
  filter(waterYear >= 1987 & waterYear <= 2021) %>% 
  mutate(residual = (all_ave_temp-aver_ann_temp)+temperature_mean-aver_day_temp) %>% 
  mutate(deviation = abs(residual-lag(residual)))
```


```{r sd2}
standard_dev_87 <- standard_dev %>% 
  filter(waterYear == 1987) %>% 
  group_by(waterYear) %>% 
  mutate(resid_mean = mean(residual)) %>%
           mutate(sd_1 = residual-resid_mean)

standard_dev_87 <- standard_dev_87 %>%
  group_by(waterYear) %>%
  mutate(sd_2 = (((sum((sd_1)^2))/((sum(tabulate(standard_dev_87$waterDay)))-1)))^(0.5))
```


```{r sd3}

standard_dev_88 <- standard_dev %>% 
  filter(waterYear == 1988) %>% 
  group_by(waterYear) %>% 
  mutate(resid_mean = mean(residual)) %>%
           mutate(sd_1 = residual-resid_mean)

standard_dev_88 <- standard_dev_88 %>%
  group_by(waterYear) %>%
  mutate(sd_2 = (((sum((sd_1)^2))/((sum(tabulate(standard_dev_87$waterDay)))-1)))^(0.5))


```


```{r sd4}

standard_dev_89 <- standard_dev %>% 
  filter(waterYear == 198) %>% 
  group_by(waterYear) %>% 
  mutate(resid_mean = mean(residual)) %>%
           mutate(sd_1 = residual-resid_mean)

standard_dev_89 <- standard_dev_89 %>%
  group_by(waterYear) %>%
  mutate(sd_2 = (((sum((sd_1)^2))/((sum(tabulate(standard_dev_87$waterDay)))-1)))^(0.5))

```

## Standard Deviation each year

```{r sd5}
standard_dev_all <- standard_dev %>% 
  group_by(waterYear) %>% 
  mutate(nmbr = n())

standard_dev_all <- standard_dev_all %>% 
  group_by(waterYear) %>% 
  mutate(resid_mean = mean(residual)) %>%
  mutate(sd_1 = residual-resid_mean) %>% 
  mutate(sd_2 = (((sum((sd_1)^2))/((nmbr-1))))^(0.5)) %>%
  distinct(sd_2, .keep_all = TRUE) %>% 
   select(waterYear, sd_2)

standard_dev_all %>% 
  kable(.,'html') %>%
  kable_styling() %>%
  scroll_box(width='250px',height='500px')

ggplot(standard_dev_all, aes(x = waterYear, y = sd_2))+#, color = waterYear)) +
  geom_line(size= 0.7) +
  #geom_line(aes) +
  theme_few() +
  geom_smooth(method = "lm", se=FALSE) +
  ylab('SD') + 
  xlab('Water year')
```

```{r sd mk & ss}

sd_mk <- mk.test(standard_dev_all$sd_2)
print(sd_mk)

sd_sens <- sens.slope(standard_dev_all$sd_2)
print(sd_sens)

```

# Notes from SF, 26/7

Need to look at seasonal variation, with examples of Nov-Mar & Jun-Aug.

In addition to Tmean, analyze Tmax & Tmin

Run seasonal MK & SS tests rather than over the whole time series. (Look at 2018 Water paper as well as 2020 Water Paper.)

*Need to correct data for SNOTEL sensor changes and figure out SNOTEL homogeneity. Obvious shift in above SD figure.* 

Stack the 2021 and 1987 figures on top of each other.

Moving to COOP data until sensor change is identified.

```{r sd6, eval=FALSE, include=FALSE}
# Trying to get the sum of waterYear occurances

# standard_dev_all <- standard_dev %>% 
#   group_by(waterYear) %>% 
#   summarise(n = sum(tabulate(standard_dev_all$waterDay)))

# standard_dev_all <- standard_dev %>% 
#   group_by(waterYear) %>% 
#   mutate(resid_mean = mean(residual)) %>%
#   mutate(sd_1 = residual-resid_mean) %>% 
#   mutate(sd_2 = (((sum((sd_1)^2))/((sum(tabulate(standard_dev_all$waterDay))))))^(0.5)) %>%
#   distinct(sd_2, .keep_all = TRUE) %>% 
#    select(waterYear, sd_2)

# ggplot(standard_dev_all, aes(sd_2)) +
#   geom_histogram(alpha = 0.5, color = "blue", bins = 6, binwidth = 1) +
#   theme_base() +
#   xlab("Temperature (°C) Above or Below Mean") +
#   ylab("Count")
#    
# ggplot(standard_dev_all, aes(sd_2)) +
#   geom_histogram(alpha = 0.5, color = "blue", bins = 6, binwidth = 1) +
#   theme_base() +
#   xlab("Temperature (°C) Above or Below Mean") +
#   ylab("Count")


```

