---
title: "CRB Variance"
author: "Brian"
date: "6/14/2022"
output:
  html_document:
    toc: true
    toc_float: true
editor_options: 
  chunk_output_type: console
---

# Disclaimer 

*This work is very preliminary as I get back into the coding swing of things. Data wrangling and figure generation will be done via R, but the rest of the project will be done using good ol' microsoft products. This is just an entry point into data crunching and should by no means be considered a final product.*

# Steamboat Tower SNOTEL

```{r libraries, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(snotelr)
library(riem)
library(dplyr)
library(ggplot2)
library(tidyverse)
library(dataRetrieval)
library(lubridate)
library(sf)
library(ggthemes)
library(xts)
library(dygraphs)
library(scales)
library(openair)
library(plotly)
library(SciViews)
knitr::opts_chunk$set(message = F, 
                      warning = F,
                      cache = T)
knitr::opts_chunk$set(echo = TRUE)
library(trend)
library(nhdplusTools)
library(lfstat)
library(ggpubr)
```

## Raw data download

### Steamboat Tower (825) SNOTEL data

```{r read in and download, echo=TRUE}
# The fucntion snotel_download evidently downloads everything in metric units. 

snotel_825_tower <- snotel_download(site_id = 825, path = tempdir('../data'), internal = TRUE)

write.csv(snotel_825_tower,"C:/Users/13074/Documents/ESS580/thesis_project/thesis_research/data_raw/snotel_825.csv", row.names = FALSE) #write in the raw data

```

### Data Cleaning

```{r clean}
head(snotel_825_tower) # check the date, usually a character.  

snotel_825_tower$Date <- as.Date(snotel_825_tower$date) #change date from character to date format, capitalize to work with Water year functon from NWIS.

snotel_825_clean <- snotel_825_tower %>% # filter for the timeframe
  filter(Date >= "1978-10-01" & Date <= "2021-08-31") %>%
  filter(temperature_mean >= -30 & temperature_mean <= 20) %>% # removing outliers   
  addWaterYear() %>% 
  mutate(daymonth = format(as.Date(Date), "%d-%m")) %>% 
  na.omit()
```

```{r 825, eval=FALSE, include=FALSE}
snotel_825$Date <- as.Date(snotel_825$date) #change date from character to date format, capitalize to work with Water year functon from NWIS.

snotel_825_clean <- snotel_825 %>% # filter for the timeframe
  filter(Date >= "1978-10-01" & Date <= "2021-08-31") %>%
  filter(temperature_mean >= -30 & temperature_mean <= 20) %>% # removing outliers   
  addWaterYear() %>% 
  mutate(daymonth = format(as.Date(Date), "%d-%m")) %>%
  mutate(DOY = yday((Date))) %>%
  na.omit()
```

##### adding water day using difftime (SUPER COOL. example from [this](https://stackoverflow.com/questions/48123049/create-day-index-based-on-water-year))

```{r clean water day}
#adding water day using difftime (SUPER COOL. example from [this](https://stackoverflow.com/questions/48123049/create-day-index-based-on-water-year))

snotel_825_clean <- snotel_825_clean %>% 
  group_by(waterYear)%>% 
  mutate(waterDay = (as.integer(difftime(Date, ymd(paste0(waterYear - 1 ,'-09-30')), units = "days"))))

```

### Figure check

```{r 825 simple plot}
ggplot(snotel_825_clean, aes(x = Date, y = temperature_mean, filter(waterYear = 1990))) +
  geom_point() + #lwd = 2) +
  theme_few() +
  geom_smooth(method = "lm", se=FALSE) +
  ylab('Daily temperature (°C)') + 
  xlab('Date')

#Check for outliers....

#dygraph

temp_xts <- xts(snotel_825_clean$temperature_mean, order.by = snotel_825_clean$Date)

dygraph(temp_xts) %>%
  dyAxis("y", label = "Daily temperature (°C)") 


```

## Detrending Data 

```{r detrending data}
#SF figured out the yearly average by water year

#average water year temperature

yearly_wy_aver <- snotel_825_clean %>% 
  group_by(waterYear) %>% 
  mutate(aver_ann_temp = mean(temperature_mean))

```


```{r detrending data2}
#Average temperature by day for all water years:

daily_wy_aver <- yearly_wy_aver %>% 
  group_by(daymonth) %>% 
  mutate(aver_day_temp = mean(aver_ann_temp))

#average mean temperature by day for the period of record:

daily_wy_aver <- daily_wy_aver %>% 
  group_by(daymonth) %>% 
  mutate(all_ave_temp = mean(daily_wy_aver$aver_day_temp))

#str(daily_wy_aver)

```

```{r detrend figures, eval=FALSE, include=FALSE}
#making day-month numeric

head(daily_wy_aver)

# try to show all years as means. Uses Jan 1 as day 1. 
daily_wy_aver2 <- daily_wy_aver %>% 
  #filter(waterYear == "1987" | waterYear == "2021") %>%
  select(DOY, daymonth, aver_day_temp) %>%
  distinct(daymonth, .keep_all = TRUE) %>% # had to use daymonth to keep it distinct
  mutate(date = as.Date(DOY)) #This worked, but gave a year of 1970?

daily_wy_aver2$aver_day_temp <- signif(daily_wy_aver2$aver_day_temp,2) #reduce the sig figs

daily_wy_aver2$DOYwy <- 1:nrow(daily_wy_aver2) #adding a number as a wateryear day of year.  

### ### ###

#NONE OF THIS WAS RIGHT- I WANTED SIMPLY THE AVERAGE TEMPERATURE BY waterDay ACROSS ALL YEARS! 
# try to show all years as means. 
daily_wy_aver2 <- daily_wy_aver %>% 
  #filter(waterYear == "1987" | waterYear == "2021") %>%
  select(waterDay, daymonth, aver_day_temp) %>%
  distinct(daymonth, .keep_all = TRUE) #%>% # had to use daymonth to keep it distinct
  #mutate(date = as.Date(DOY)) #This worked, but gave a year of 1970?
  #NO NEED TO DO THIS, HAVE waterDay NOW.

daily_wy_aver2$aver_day_temp <- signif(daily_wy_aver2$aver_day_temp,2) #reduce the sig figs



```

```{r Figure all year average temp}

# try to show all years as means. 
daily_wy_aver2 <- daily_wy_aver %>% 
  #filter(waterYear == "1987" | waterYear == "2021") %>%
  group_by(waterDay) %>%
  mutate(date_temp = mean(temperature_mean))
  

daily_wy_aver2$date_temp <- signif(daily_wy_aver2$date_temp,2) #reduce the sig figs



ggplot(daily_wy_aver2, aes(x = waterDay, y = date_temp))+#, color = waterYear)) +
  geom_line(size= 0.7) +
  #geom_line(aes) +
  theme_few() +
  #geom_smooth(method = "lm", se=FALSE) +
  ylab('Average Daily temperature (°C)') + 
  xlab('Day of water year')
```
**Day of year average temperature for the 1987-2021 period of record for Tower SNOTEL site.**


## All years vs 1987 & 2021

```{r all year average join to 1987 & 2021 WY}



# 87 and 21 only. Not useful
daily_wy_aver3 <- daily_wy_aver2 %>% 
  filter(waterYear == "1987" | waterYear == "2021") %>%
  select(waterDay, waterYear, temperature_mean, daymonth, date_temp) %>%
  group_by(daymonth)

# ggplot(daily_wy_aver3, aes(x = waterDay, y = date_temp))+#, color = waterYear)) +
#   geom_line(size= 0.7) +
#   geom_line(aes(y = temperature_mean, color = waterYear)) +
#   theme_few() +
#   #geom_smooth(method = "lm", se=FALSE) +
#   #scale_colour_identity() +
#   scale_color_manual(name = "Water Year", values = c("blue", "red")) 
#   ylab('Daily temperature (°C)') + 
#   xlab('Day of water year')

ggplot(daily_wy_aver3) + #, color = waterYear)) +
  geom_line(aes(x = waterDay, y = date_temp), group = 1, size= 0.7) +
  geom_line(aes(x = waterDay, y = temperature_mean, group = 1, colour = waterYear)) +
  theme_few() +
  #geom_smooth(method = "lm", se=FALSE) +
  #scale_colour_identity() +
  
  #scale_color_manual(name = "Water Year") 
  ylab('Daily temperature (°C)') + 
  xlab('Day of water year')
  
# ??geom_line  
#   
# scale_color_manual(name = "Colors", 
#                      values = c("a" = "blue", "b" = "red")  
#   
# color_group <- c("blue","black")
#   
# ?scale_discrete_identity()

```
**Day of year average temperature and daily 1987 & 2021 temperatures for the 1987-2021 period of record for Tower SNOTEL site.**


```{r factor day labels, eval=FALSE, include=FALSE}

daily_wy_aver2$DOY <- factor(daily_wy_aver2$DOY, levels=c(274, 275, 276, 277, 278 279 280 281 282 283 284 285 286 287 288 289 290 291 292 293 294 295 296 297 298 299
[27] 300 301 302 303 304 305 306 307 308 309 310 311 312 313 314 315 316 317 318 319 320 321 322 323 324 325
[53] 326 327 328 329 330 331 332 333 334 335 336 337 338 339 340 341 342 343 344 345 346 347 348 349 350 351
[79] 352 353 354 355 356 357 358 359 360 361 362 363 364 365 366, 1, 2, 3, 4:273))

?factor

#Check for outliers....

#dygraph

temp_xts <- xts(snotel_825_clean$temperature_mean, order.by = snotel_825_clean$Date)

dygraph(temp_xts) %>%
  dyAxis("y", label = "Daily temperature (°C)") 

#EDIT. The above didn't work, trying a different approach. 

```


### Calculating varience

```{r Calculating residuals}
#Calculating residuals, WY 1987 (first full year on record for 825)

residuals_1987 <- daily_wy_aver %>% 
  group_by(waterYear) %>% 
  filter(waterYear == 1987) %>% 
  mutate(residual = (aver_ann_temp-all_ave_temp)+temperature_mean-aver_day_temp) %>% 
  mutate(deviation = abs(residual-lag(residual)))

residuals_2021 <- daily_wy_aver %>% 
  group_by(waterYear) %>% 
  filter(waterYear == 2021) %>% 
  mutate(residual = (aver_ann_temp-all_ave_temp)+temperature_mean-aver_day_temp) %>% 
  mutate(deviation = abs(residual-lag(residual)))


```

## Figures 

### 1987

```{r residuals 1987}

ggplot(residuals_1987, aes(residual)) +
  geom_histogram(alpha = 0.5, color = "blue", bins = 15) +
  theme_base() +
  xlab("Temperature (°C) Above or Below Mean") +
  ylab("Count")

```

**Distribution of daily temperature variance for 1987 compared to the 1987-2021 period of record.**


```{r deviations 1987}

ggplot(residuals_1987, aes(deviation)) +
  geom_histogram(alpha = 0.5, color = "blue", bins = 15) +
  theme_base() +
  xlab("Diurnal Temperature Fluctuation (°C)") +
  ylab("Count")

```

**Distribution of 1987 diurnal temperature fluctuation from the 1987-2021 period of record.**

### 2021

```{r residuals 2021}

ggplot(residuals_2021, aes(residual)) +
  geom_histogram(alpha = 0.5, color = "blue", bins = 15) +
  theme_base() +
  xlab("Temperature (°C) Above or Below Mean") +
  ylab("Count")

```

**Distribution of daily temperature variance for 2021 compared to the 1987-2021 period of record.**

```{r deviations 2021}

ggplot(residuals_2021, aes(deviation)) +
  geom_histogram(alpha = 0.5, color = "blue", bins = 15) +
  theme_base() +
  xlab("Diurnal Temperature Fluctuation (°C)") +
  ylab("Count")

```

**Distribution of 2021 diurnal temperature fluctuation from the 1987-2021 period of record.**

# Next steps / Further questions

### SNOTEL

Need to investigate site specific issues related to disturbances (general SNOTEL issues). 

Having issues with day-of-year and water year interactions. This keeps me from showing mean trendlines over the whole period of record as R does not want to recognize the x axis as anything other than numerical.   

Would using a pivot table be better when displaying temperature for 240 sites? (What data management is best for that much data?)

How to include oscillation influence into the data? 

Need to include a couple statistic tests. 

### Beyond SNOTEL

What other data / site types do I want to include? 

### Coding

I am sure Matt would tell me to automate this. Not quite sure how, but it seems likely I could find a list of SNOTEL sites and using spatial data, exclude sites not within the CRB boundary. From there, I could write a for loop & function to use the snotelr function to download data. *Jack Reuland helped with this idea below. Now what to do with the large dataframe....*

# All stations:

Per Jack's recommendations, BUT, while there are 240 stations, I could only find 166 while clipping the NRCS SNOTEL points and a CRB polygon from CAPGISadmin. The NRCS stated the sites given were active, so perhaps there are 74 inactive sites?

```{r read in SNOTEL station site_ids from ArcGIS, echo=TRUE}

TableOfSNOTEL_StudyArea <- read.csv("C:/Users/13074/Desktop/MS_Watershed_Science/Literature/SNOTEL_sites_1_2.csv", header = TRUE)

studyarea_site_id <- as.vector(TableOfSNOTEL_StudyArea$site_id)

SNOTEL_StudyArea <- snotel_download(site_id = c(studyarea_site_id), internal = TRUE)

write.csv(SNOTEL_StudyArea,"C:/Users/13074/Documents/ESS580/thesis_project/thesis_research/data_raw/SNOTEL_StudyArea.csv", row.names = FALSE) #write in the raw data

```

```{r summary StudyArea}

summary(SNOTEL_StudyArea)

#lol, this is meaningless. 
```

# Notes from SF, 22/6:

Some initial thoughts: 1) I assume that frequency is the raw count, i.e., number of days? At some point, a % of the year may be better; 2) COOP data, such as at Steamboat, should have a much longer period of record if you want to examine change over time - setting up the analysis is a good start; 3) there is an inconsistency with the SNOTEL data (see https://doi.org/10.1029/2019WR025921); 4) we (https://doi.org/10.1029/2002WR001512, etc.) used 240 stations (metadata at https://doi.org/10.1029/2009WR007835), and I know that are a few are inactive, but likely not 74 - some are outside of the basin (https://agupubs.onlinelibrary.wiley.com/cms/asset/2af0fc20-a948-455d-bf0e-8f6e7dcc7ad9/wrcr12217-fig-0001.png) but still representative (a few can be removed, in hindsight, like the ones in the Laramie Range).

# Notes, Monday 27/6

Figured out the water day issue and applied it to the Day of water year figures (this was an issue that I spent waaaaaay too much time on). Went the wrong way with pivot tables, day of year, and using average annual temp for the mean aily figure.  Going to use Excel to check if correct. Goal for Tuesday: Determine daily percentages.